
---
output: # rmarkdown::github_document
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 4 - Extra Credits"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment4_extra_credits.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- Notice that if there is `eval=FALSE` in the first line of the code chunk, the chunk will not be execute.

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

-------

This assignment works with the [IMDB Top 1000 data](../data/imdb_top_1000.csv).  Find out more information about this data at [this link](https://www.kaggle.com/datasets/harshitshankhdhar/imdb-dataset-of-top-1000-movies-and-tv-shows).  Import the data and answer the following questions. 


```{r}
library(readr)
df <- read_csv("C:\\Users\\student\\OneDrive - Bryant University\\MATH421\\GitHub\\MATH421ML\\MATH421ML\\imdb_top_1000.csv")
```

1. List all the names of the columns of the data
```{r}
names(df)
```


2. Which movies have the highest money earned (Gross)?
```{r}
library(dplyr)
df %>%
  arrange(desc(Gross)) %>%
  select(Series_Title, Gross)
```



3. What is the lowest rating (IMDB_Rating)?  List five movies have this lowest rating.

```{r}
df %>%
  filter(IMDB_Rating == min(IMDB_Rating, na.rm = TRUE)) %>%
  select(Series_Title, IMDB_Rating)
```


4. Which year have the most number of movies released in the list?  What is the total of money earned on that year?

```{r}
library(dplyr)

df %>%
  group_by(Released_Year) %>%
  summarise(
    Number_of_Movies = n(),
    Total_Gross = sum(Gross, na.rm = TRUE)
  ) %>%
  arrange(desc(Number_of_Movies)) %>%
  head(1)
```



5. What is the total money earned per movies on average?

```{r}
library(dplyr)
df %>%
  summarise(Average_Gross = mean(Gross, na.rm = TRUE))
```


6. Calculate the average number of votes by year. Calculate the average number of votes of movies that have IMDB rating greater than 9.

```{r}
library(dplyr)
# Part A 
df %>%
  group_by(Released_Year) %>%
  summarise(Average_Votes = mean(No_of_Votes, na.rm = TRUE))

# Part B 
df %>%
  filter(IMDB_Rating > 9) %>%
  summarise(Average_Votes = mean(No_of_Votes, na.rm = TRUE))
```


7. Calculate the average Meta score in 2020 of movies that have number of votes in the third quartile. 

```{r}
library(dplyr)
#if this question is indicating that it is above the third quartile 
q3 <- quantile(df$No_of_Votes, 0.75, na.rm = TRUE)
df %>%
  filter(Released_Year == 2020,No_of_Votes >= q3) %>%
  summarise(Average_Meta_Score = mean(Meta_score, na.rm = TRUE))

#if this means that it is in between 50 and 75 
q2 <- quantile(df$No_of_Votes, 0.50, na.rm = TRUE)
q3 <- quantile(df$No_of_Votes, 0.75, na.rm = TRUE)
df %>%
  filter( Released_Year == 2020, No_of_Votes >= q2, No_of_Votes < q3) %>%
  summarise(Average_Meta_Score = mean(Meta_score, na.rm = TRUE))


```



8. (Optional - Challenging). The current `Runtime` variable is not a numeric. Use the `str_remove` function to remove ` min` from the variables then use `as.numeric` to convert the variable to numeric. Calculate the average running time in the 2010s. Calculate the correlation between running time and rating (adding `use="complete.obs"` in the `cor` function to ignore the missing values).  

9. We can use `select_if` to select columns satisfying a condition and use `summarise_if` to do calculation on columns satisfying a condition.  Try the follows to understand these functions. 

```{r,eval=TRUE}
# Select only character columns
df %>% select_if(is.character)

# Calculate the median of all numeric columns
df %>% summarise_if(is.numeric, mean, na.rm=TRUE)
```


10. Implement the follows functions or combos. Drawing a comment or summary from each calculation. The codes in this question should be different from the codes used in other questions.

- select
- filter
- mutate
- summarise
- arrange
- count
- count + arrange
- filter + count + arrange
- group_by + summarise
- filter + group_by + summarise
- filter + group_by + summarise + arrange

```{r}
#select
df %>%
select(Series_Title,Runtime)
#filter
df %>%
  filter(Meta_score == min(Meta_score, na.rm = TRUE)) %>%
  select(Series_Title)
#mutate - giving rating out of 100
df %>%
mutate(Rating100 = IMDB_Rating * 10)
#summarise
df %>%
summarise(avg_rating = mean(IMDB_Rating, na.rm = TRUE))
#arrange sorting by rating 
df %>%
  arrange(desc(IMDB_Rating))

#count counting how many have each certificate 
df %>%
  count(Certificate)

#count + arrange arranging what i just counted in decreasing order of the count 
df %>%
  count(Certificate) %>%
  arrange(desc(n))

# filter + count + arrange added a filter for the rating being higher than 8.5 
df %>%
  filter(IMDB_Rating > 8.5) %>%
  count(Certificate) %>%
  arrange(desc(n))


# group by and summarize average rating by each certificate group 
df %>%
  group_by(Certificate) %>%
  summarise(avg_rating = mean(IMDB_Rating, na.rm = TRUE))


# filter group by summazire - added filter for no of votes to be greater than 100000
df %>%
  filter(No_of_Votes > 100000) %>%
  group_by(Certificate) %>%
  summarise(avg_rating = mean(IMDB_Rating, na.rm = TRUE))

# filter + group_by + summarise + arrange - average meta score when the imdb rating it greater than 8.5.grouped by certificate and then organized into decreasing order 
df %>%
  filter(IMDB_Rating > 8.5) %>%
  group_by(Certificate) %>%
  summarise(avg_meta = mean(Meta_score, na.rm = TRUE)) %>%
  arrange(desc(avg_meta))


```






