---
title: "Assignment Two"
author: "Megan Luby"
format: html
editor: visual
---

***How to do it?***:

-   Open the Quarto file of this assignment ([link](assignment2q.qmd)) in Rstudio.

-   Right under each **question**, insert a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question.

-   `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.

-   Publish the html file to your Githiub Page.

***Submission***: Submit the link on Github of the assignment to Canvas.

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```

------------------------------------------------------------------------

## Problems

1.  Calculate:

$$
S = 1 + 4 + 7 + 10 +... + 100
$$
```{r}
x <- seq(1,100,3)
sum(x)
```

2.  Calculate:

$$
S = 1^{100} + 2^{100} + 3^{100} +...+100^{100}
$$
```{r}
sum((1:100)^100)
```

3.  Calculate:

$$
S = \frac{1}{1\cdot 2}+\frac{1}{2 \cdot 3}+...+\frac{1}{100\cdot 101}
$$
```{r}
n <- c(1:100)
sum(1/ (n * (n + 1)))
```

4.  Calculate:

$$
S = \frac{1}{1\cdot 2 \cdot 3}+\frac{1}{2 \cdot 3 \cdot 4}+...+\frac{1}{100\cdot 101 \cdot 102}
$$
```{r}
n <- (1:100)
sum(1 / (n * (n + 1) * (n + 2)))
```
5.  Calculate:

$$
S = 1^1 + 3^3 + 5^5 + 7^7 +....+101^{101}

$$
```{r}
n <- seq(1,101,2)
sum(n^n)
```
