
---
output: 
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 8: Visualization - Making a bar race"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment8.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```


-------

1. Install two packages `gganimate` and `gifski` then restart Rstudio. Using the `Adult Census Income` data, make an animation using `geom_point` and `transition_states`.

```{r}
library(readr)
df <- read_csv("https://bryantstats.github.io/math421/data/adult_census.csv")


library(gganimate)
library(ggplot2)
library(tidyverse)
library(knitr)
df %>%
  ggplot(aes(x = age, y = capital.gain)) +
  geom_point() +
  transition_states(marital.status) +
  labs(
    x = "Age",
    y = "Capital Gain",
    title = "Marital Status: {closest_state}"
  )
```


2. Using the `Adult Census Income` data, make an animation using `geom_bar` and `transition_states`.

```{r}
library(readr)
df <- read_csv("https://bryantstats.github.io/math421/data/adult_census.csv")


library(gganimate)
library(ggplot2)
library(tidyverse)
library(knitr)
df %>%
  ggplot(aes(x = sex)) +
  geom_bar() +
  transition_states(marital.status) +
  labs(
    x = "Sex",
    y = "Count",
    title = "Marital Status: {closest_state}"
  )

```


3. Use the WHO's dataset at [this link](https://covid19.who.int/WHO-COVID-19-global-data.csv). Make a top-10 bar race by months between countries on the number of deaths by Covid 19 in 2021.

```{r}
library(readr)
df1 <- read_csv('https://bryantstats.github.io/math421/data/WHO-COVID-19-global-data.csv')
library(tidyverse)
library(lubridate)
library(gganimate)

# Keep only 2021 and create month variable
df1 <- df1 %>%
  mutate(Date_reported = as.Date(Date_reported)) %>%
  filter(year(Date_reported) == 2021) %>%
  mutate(month = month(Date_reported, label = TRUE))

# Total deaths by country and month
d1 <- df1 %>%
  group_by(month, Country) %>%
  summarise(deaths = sum(New_deaths), .groups = "drop")

# Rank countries within each month
d2 <- d1 %>%
  group_by(month) %>%
  mutate(rank = rank(-deaths))

# Keep top 10 countries
d3 <- d2 %>%
  filter(rank <= 10)

# Bar race
a1 <- d3 %>%
  ggplot(aes(x = rank,
             y = deaths,
             group = Country,
             fill = Country,
             label = Country)) +
  geom_col() +
  geom_text(aes(label = Country), hjust = 1.2) +
  coord_flip(clip = "off", expand = FALSE) +
  scale_x_reverse() +
  labs(
    title = "Month: {closest_state}",
    x = "",
    y = "Total COVID Deaths",
    fill = "Country"
  ) +
  theme(
    plot.title = element_text(hjust = 1, size = 22),
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank()
  ) +
  transition_states(month) +
  ease_aes("cubic-in-out")

animate(a1, nframes = 400)

```


4. Make a bar race using a dataset of your own interest. You may use the dataset that we use in class (https://covidtracking.com/data/download/all-states-history.csv) but you should make a different bar racev from ones in the slides. 


```{r}
library(readr)
df2 <- read_csv("https://covidtracking.com/data/download/all-states-history.csv")


library(tidyverse)
library(lubridate)
library(gganimate)

df2 <- df2 %>%
  mutate(
    date = as.Date(date),
    month = month(date, label = TRUE)
  )

d1 <- df2 %>%
  group_by(month, state) %>%
  summarise(
    deaths = sum(deathIncrease, na.rm = TRUE),
    .groups = "drop"
  )

d2 <- d1 %>%
  group_by(month) %>%
  mutate(rank = rank(-deaths))

d3 <- d2 %>%
  filter(rank <= 10)

a1 <- d3 %>%
  ggplot(
    aes(
      x = rank,
      y = deaths,
      group = state,
      fill = state,
      label = state
    )
  ) +
  geom_col() +
  geom_text(aes(label = state), hjust = 1.4) +
  coord_flip(clip = "off", expand = FALSE) +
  scale_x_reverse() +
  labs(
    title = "Month: {closest_state}",
    x = "",
    y = "Total Death Increases",
    fill = "State"
  ) +
  theme(
    plot.title = element_text(hjust = 1, size = 22),
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank()
  ) +
  transition_states(month) +
  ease_aes("cubic-in-out")

animate(a1, nframes = 400)


```




