
---
output: 
  html_document:
  pdf_document: default
  word_document: default
title: "Assignment 9: Visualization - Revealing a graph"
---

***How to do it?***: 

- Open the Rmarkdown file of this assignment ([link](assignment9.Rmd)) in Rstudio. 

- Right under each **question**, insert  a code chunk (you can use the hotkey `Ctrl + Alt + I` to add a code chunk) and code the solution for the question. 

- `Knit` the rmarkdown file (hotkey: `Ctrl + Alt + K`) to export an html.  

-  Publish the html file to your Githiub Page. 

***Submission***: Submit the link on Github of the assignment to Canvas

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```


-------

**Notice:** *In this assignment, all the plot should have title, caption, and axis labels. *

1. Use the dataset by of covid 19 by WHO at https://covid19.who.int/WHO-COVID-19-global-data.csv. Find the three countries with the most numbers of deaths by Covid-19. 

Hint: 

-  Use `filter` to select the most current date 


-  Pipe with `arrange(-Cumulative_deaths)` to rank the data by the number of cumulative deaths

```{r}
library(readr)
df <- read_csv("https://bryantstats.github.io/math421/data/WHO-COVID-19-global-data.csv")

library(tidyverse)


df %>%
  filter(Date_reported == max(Date_reported)) %>%
  arrange(desc(Cumulative_deaths)) %>%
  select(Country, Cumulative_deaths) %>%
  head(3)

```




2. Use `transition_reveal` to make a plot revealing the number of deaths in the three countries with the most numbers of deaths


```{r}
library(tidyverse)
library(gganimate)

top3 <- df %>%
  filter(Date_reported == max(Date_reported)) %>%
  arrange(desc(Cumulative_deaths)) %>%
  slice(1:3) %>%
  pull(Country)

p <- df %>%
  filter(Country %in% top3) %>%
  mutate(Date_reported = as.Date(Date_reported)) %>%
  ggplot(
    aes(
      x = Date_reported,
      y = Cumulative_deaths,
      color = Country
    )
  ) +
  geom_line(size = 1) +
  labs(
    title = "COVID-19 Cumulative Deaths",
    subtitle = "{frame_along}",
    x = "Date",
    y = "Cumulative Deaths"
  ) +
  transition_reveal(Date_reported)

animate(p, nframes = 300)
```


3. Create the new variable (column) `death_per_cases` recording the number of deaths per cases (Hint: divide cumulative deaths by cumulative cases).  What are the three countries with the highest deaths per cases? 

```{r}
df %>%
  filter(Date_reported == max(Date_reported)) %>%
  mutate(death_per_cases = Cumulative_deaths / Cumulative_cases) %>%
  arrange(desc(death_per_cases)) %>%
  select(Country, death_per_cases) %>%
  head(3)
```



4. Use `transition_reveal` to make a plot revealing the number of deaths per cases of the US, Italy and Mexico. 

```{r}
library(tidyverse)
library(gganimate)

three <- df %>%
  filter(Country %in% c("United States of America", "Italy", "Mexico")) %>%
  mutate(
    Date_reported = as.Date(Date_reported),
    death_per_cases = Cumulative_deaths / Cumulative_cases
  ) %>%
  ggplot(
    aes(
      x = Date_reported,
      y = death_per_cases,
      color = Country
    )
  ) +
  geom_line(size = 1) +
  labs(
    title = "Deaths per Cases by Country",
    subtitle = "{frame_along}",
    x = "Date",
    y = "Deaths per Cases"
  ) +
  transition_reveal(Date_reported)

animate(three, nframes = 300)
```

5. Import the following data https://raw.githubusercontent.com/owid/covid-19-data/master/public/data/vaccinations/vaccinations.csv
                      
Use `transition_reveal` to plot the total vaccinations per hundred by level of income. Hint: the levels of income are identified in the `location` variable. Put a short comment on the caption of the plot. 


```{r}
df1 <- read_csv("https://raw.githubusercontent.com/owid/covid-19-data/master/public/data/vaccinations/vaccinations.csv")
library(tidyverse)
library(gganimate)

income <- df1 %>%
  filter(location %in% c(
    "Low income",
    "Lower middle income",
    "Upper middle income",
    "High income"
  )) %>%
  mutate(date = as.Date(date)) %>%
  ggplot(
    aes(
      x = date,
      y = total_vaccinations_per_hundred,
      color = location
    )
  ) +
  geom_line(size = 1) +
  labs(
    title = "Total Vaccinations per Hundred by Income Level",
    subtitle = "{frame_along}",
    x = "Date",
    y = "Total Vaccinations per Hundred",
    caption = "Higher-income groups generally reached higher vaccination levels faster than lower-income groups."
  ) +
  transition_reveal(date)

animate(income, nframes = 300)
```

